# Quick start on taxlist
*Miguel Alvarez*

<br>

## Introduction
This is a guide for constructing `taxlist` objects via command lines.
For accessing this document from your **R console** use following command:

```{r, eval=FALSE}
browseURL("https://github.com/kamapu/Guides/blob/master/taxlist_firststeps.md")
```

To install the last version of `taxlist`, you can use the package `devtools`:

```{r,eval=FALSE}
library(devtools)

install_github("kamapu/taxlist")
```

## Starting with building blocks

I will take an example from "Helechos de Chile" **(Gunkel 1984)** to demonstrate
how to construct a `taxlist` object from its building blocks.
The first step will be to generate an empty `taxlist` object:

```{r}
library(taxlist)

Fern <- new("taxlist")
summary(Fern)
```

As you can see, there is nothing in there.
We start including taxonomic levels, we like to insert in the list.
Remember, the levels have to be provided in an upward sequence, that is to say
from lower to higher levels:

```{r}
levels(Fern) <- c("variety","species","genus")
```
Then you can add a species:

```{r}
Fern <- add_concept(Fern, TaxonName="Asplenium obliquum", AuthorName="Forster",
	Level="species")
```

Then add varieties:

```{r}
Fern <- add_concept(Fern,
	TaxonName=c("Asplenium obliquum var. sphenoides",
		"Asplenium obliquum var. chondrophyllum"),
	AuthorName=c("(Kunze) Espinosa",
		"(Bertero apud Colla) C. Christense & C. Skottsberg"),
	Level="variety")
```

Finally add the genus and check the object:

```{r}
Fern <- add_concept(Fern, TaxonName="Asplenium", AuthorName="L.", Level="genus")
summary(Fern)

summary(Fern, "all")
```

## Set parent-child relationships and synonyms

Now set the parent-child relations.
Relating to the previous display, you know that the species (concept ID **1**)
is the parent of the varieties (IDs **2** and **3**), and the genus (ID **4**)
is the parent of the species (ID **1**).
Thus the relationships are set as:

```{r}
add_parent(Fern, c(2,3)) <- 1
add_parent(Fern, 1) <- 4
summary(Fern)
```

Similarly to the addition of concepts, you can also add synonyms:

```{r}
Fern <- add_synonym(Fern, ConceptID=2, TaxonName=c("Asplenium sphenoides"),
	AuthorName="Kunze")
summary(Fern, "all")
```

Hierarchical levels, parent-child relationships and synonyms are included in the
exemplary data `Easplist`.
For further functions, look to the package's manual.
